\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-06-07 nampi.tex
shortened

Much of everyday human thinking and learning can be understood in
terms of program induction: constructing a procedure that maps inputs
to desired outputs, based on observing example input-output pairs.
People can induce programs flexibly across many different domains, and
remarkably, often from just one or a few examples.  For instance, if
shown that a text-editing program should map ``Jane Morris Goodall''
to ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey''
to ``R. E. Leakey''; if instead the first input mapped to
``Dr. Jane'', ``Goodall, Jane'', or ''Morris'', we might have guessed
the latter should map to ``Dr. Richard'', ``Leakey, Richard'', or
``Erskine'', respectively.

The FlashFill system~\cite{gulwani2011automating} embedded in
Microsoft Excel solves problems such as these and is probably the best
known practical program-induction algorithm, but researchers in
programming languages and AI have built successful program induction
algorithms for many applications, such as handwriting recognition and
generation~\cite{lake2015human}, procedural
graphics~\cite{ellis2017learning}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}.  These systems
work in different ways, but most hinge upon a carefully engineered
\textbf{Domain Specific Language (DSL)}.  This is especially true for
systems such as FlashFill that aim to induce a wide range of programs
very quickly, in a few seconds or less.  DSLs constrain the search
over programs with strong prior knowledge in the form of a restricted
set of programming primitives tuned to the needs of the domain: for
text editing, these are operations like appending and
splitting on characters.

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
\end{table*}
Because any computable learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on the kinds of problems that people solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  %% Even with a good DSL, program search may be
%% intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
%% so we train a
%% neural network to guide the search procedure.

@@

Our algorithm is called \system because it is based on a
novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
between ``wake'' and ``sleep'' phases to achieve three goals: finding
programs that solve tasks; creating a DSL by discovering
and reusing domain-specific subroutines; and training a neural network
that efficiently guides search for programs in the DSL.
The learned DSL effectively encodes a prior on programs likely to
solve tasks in the domain, while the neural net looks at the example
input-output pairs for a specific task and produces a ``posterior''
for programs likely to solve that specific task.  The neural network
thus functions as a \textbf{recognition model} supporting a form of
approximate Bayesian program induction, jointly trained with a
\textbf{generative model} for programs encoded in the DSL, in the
spirit of the Helmholtz machine~\cite{hinton1995wake}. The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.

@@

\subsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{lau2001programming},
and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
Here we will instead start out with generic sequence manipulation
primitives and recover many of the higher-level building blocks that have
made these other text editing systems successful.
%% Because our enumerative search procedure cannot generate string %
%% constants, we instead enumerate programs with string-valued
%% parameters.  For example, to learn a program that prepends ``Dr.'', we
%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
%% define $\probability[x|p]$ by approximately marginalizing out the
%% string parameters via a simple dynamic program.
%% In Sec.~\ref{regressionSection}, we will use a similar trick to
%% synthesize programs containing real numbers, but using gradient
%% descent instead of dynamic programming.

@@

To be clear, we believe that program learners should \emph{not}
start from scratch,
but instead should start from
a rich, domain-agnostic
basis like those embodied in the standard libraries of modern programming languages.
What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
and can in principle recover many of the amenities of modern programming systems,
provided it is given enough computational power and a suitable
spectrum of programming exercises.

@@

\section{Quantitative Results}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them~\ref{learningCurves}.
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models that
more closely match the domains.

>>>>

Much of everyday human thinking and learning can be understood in
terms of program induction: constructing a procedure that maps inputs
to desired outputs, based on observing example input-output pairs.
People can induce programs flexibly across many different domains, often from just one or a few examples.  For instance, if
shown that a text-editing program should map ``Jane Morris Goodall''
to ``J. M. Goodall'', we can guess it maps ``Richard Erskine Leakey''
to ``R. E. Leakey''; if instead the first input mapped to
``Dr. Jane'' or ``Goodall, Jane'', we might guess
the latter should map to ``Dr. Richard'' or ``Leakey, Richard'',
 respectively.

The FlashFill system~\cite{gulwani2011automating} embedded in
Microsoft Excel solves problems such as these and is probably the best
known practical program-induction algorithm, but researchers in
programming languages and AI have had successes in many domains, such as handwriting recognition and
generation~\cite{lake2015human}, procedural
graphics~\cite{ellis2017learning}, question
answering~\cite{johnson2017clevr} and robot motion
planning~\cite{devlin2017neural}.  These systems
work in different ways, but most hinge upon a carefully engineered
\textbf{Domain Specific Language (DSL)}.  This is especially true for
systems like FlashFill that induce a wide range of programs
very quickly, in a few seconds or less.  DSLs constrain the search
over programs with strong prior knowledge in the form of a restricted
set of programming primitives finely tuned to the domain: for
text editing, these are operations like appending and
splitting on characters.

@@

  \end{tabular}
  \\\bottomrule\\
\end{tabular}}
\caption{Top: Tasks from three domains we apply our algorithm to, each followed by the programs \system discovers for them. Bottom: Several examples from learned DSL. Notice that learned DSL primitives can call each other, and that \system rediscovers higher-order functions like \code{filter} ($f_1$ under List Functions)}\label{initialExampleDSL}
\end{table*}
Because any learning problem can in principle be cast as
program induction, it is important to delimit our focus.  In contrast
to computer assisted programming~\cite{solar2008program} or genetic
programming~\cite{DBLP:books/daglib/0070933}, our goal is not to automate software
engineering, or to synthesize large bodies of code starting from scratch.  Ours is a basic AI goal:
capturing the human ability to learn to think flexibly and efficiently
in new domains --- to learn what you need to know about a domain so you
don't have to solve new problems starting from scratch.  We are
focused on problems that people solve relatively
quickly, once they acquire the relevant domain expertise.  These
correspond to tasks solved by short programs --- if you have an
expressive DSL.  %% Even with a good DSL, program search may be
%% intractable;
%
% , and adding new library routines to the DSL only broadens the
% search space;
%
%% so we train a
%% neural network to guide the search procedure.

@@

Our algorithm is called \system because it is based on a
novel kind of ``wake-sleep'' learning (c.f. \cite{hinton1995wake}), iterating
between ``wake'' and ``sleep'' phases to achieve three goals: finding
programs that solve tasks; creating a DSL by discovering
and reusing domain-specific subroutines; and training a neural network
that efficiently guides search for programs in the DSL.
The learned DSL effectively encodes a prior on programs likely to
solve tasks in the domain, while the neural net looks at the example
input-output pairs for a specific task and produces a ``posterior''
for programs likely to solve that specific task.  The neural network
thus functions as a \textbf{recognition model} supporting a form of
approximate Bayesian program induction, jointly trained with a
\textbf{generative model} for programs encoded in the DSL, in the
spirit of the Helmholtz machine~\cite{hinton1995wake}. The
recognition model ensures that searching for programs remains
tractable even as the DSL %% (and hence the search space for programs)
expands.

@@

\subsection{Text Editing}\label{textSection}
Synthesizing programs that edit text is a classic problem in the
programming languages and AI literatures~\cite{lau2001programming},
and algorithms that learn text editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a hand-engineered DSL.
Here we will instead start out with generic sequence manipulation
primitives.%%  and recover many of the higher-level building blocks that have
%% made these other text editing systems successful.
%% Because our enumerative search procedure cannot generate string %
%% constants, we instead enumerate programs with string-valued
%% parameters.  For example, to learn a program that prepends ``Dr.'', we
%% enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
%% learned appending primitive (Fig.~\ref{initialExampleDSL}) --- and then
%% define $\probability[x|p]$ by approximately marginalizing out the
%% string parameters via a simple dynamic program.
%% In Sec.~\ref{regressionSection}, we will use a similar trick to
%% synthesize programs containing real numbers, but using gradient
%% descent instead of dynamic programming.

@@

We believe that program learners should \emph{not}
start from scratch,
but instead should start from
a rich, domain-agnostic
basis like those embodied in the standard libraries of modern  languages.
What this experiment shows is that \system doesn't \emph{need} to start from a rich basis,
and can in principle recover many of the amenities of modern programming systems,
provided it is given enough computational power and a suitable
spectrum of tasks.% programming exercises.

@@

\section{Quantitative Results}\label{quantitative}
We evaluate  on held-out testing tasks,
measuring how many
tasks are solved and how long it takes to solve them (Fig.~\ref{learningCurves}).
Prior to any learning,
the system cannot find solutions for most of the tasks,
and those it does solve take a long time;
with more wake/sleep iterations,
we converge upon DSLs and recognition models 
more closely matching the domain.
